Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]

### 2021-06-23 paper.Rmd
Changed data from mtcars to swiss.

---
title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]"
shorttitle: "Writing a reproducible paper with R Markdown and Pagedown"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <camille.landesvatter@mzes.uni-mannheim.de>
date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
download-link: 'Download: https://osf.io/k8jhx'
abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing). Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper". We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and R pagedown. We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
# at least one keyword must be supplied
keywords: [open science, transparency, replication, R, markdown, pagedown]
links-to-footnotes: false
paged-footnotes: true
bibliography: references.bib
output: 
  pagedown::jss_paged:
    template: wp_paged.html
    self_contained: true
    css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    csl: american-sociological-association.csl
# uncomment the next line if you only want to knit to HTML and not PDF
#knit: pagedown::chrome_print
---

@@

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown. In our view, Pagedown and the underlying technology may replace Latex in the long run.

@@

Below we import an exemplary dataset ([download](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing)).

@@

```{r}
dput(data)
```

@@

```{r eval=FALSE, include=TRUE}
data <- structure(list(X = 1:50, speed = c(4L, 4L, 7L, 7L, 8L, 9L, 10L, 
10L, 10L, 11L, 11L, 12L, 12L, 12L, 12L, 13L, 13L, 13L, 13L, 14L, 
14L, 14L, 14L, 15L, 15L, 15L, 16L, 16L, 17L, 17L, 17L, 18L, 18L, 
18L, 18L, 19L, 19L, 19L, 20L, 20L, 20L, 20L, 20L, 22L, 23L, 24L, 
24L, 24L, 24L, 25L), dist = c(2L, 10L, 4L, 22L, 16L, 10L, 18L, 
26L, 34L, 17L, 28L, 14L, 20L, 24L, 28L, 26L, 34L, 34L, 46L, 26L, 
36L, 60L, 80L, 20L, 26L, 54L, 32L, 40L, 32L, 40L, 50L, 42L, 56L, 
76L, 84L, 36L, 46L, 68L, 32L, 48L, 52L, 56L, 64L, 66L, 54L, 70L, 
92L, 93L, 120L, 85L)), 
class = "data.frame", row.names = c(NA, 
-50L))
```

@@

Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.

@@

kable(mtcars[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)

@@

```{r table-3, echo=TRUE, message=FALSE, warning=FALSE}
library(modelsummary)
datasummary_skim(mtcars, 
                 type="numeric", 
                 histogram=T, 
                 title = "Summary: Numeric variables")
```

@@

```{r table-4, echo=TRUE, message=FALSE, warning=FALSE}
# Create categorical variables
mtcars$vs_cat <- as.logical(mtcars$vs)
mtcars$cyl_cat <- as.factor(mtcars$cyl)

library(flextable)
tab_cat <- datasummary_skim(mtcars, 
                            type="categorical", 
                            title = "Summary: Categorical variables",
                            output = 'flextable')

@@

```{r fig-1, fig.align="center", fig.cap="Scatterplot of Speed and Distance", fig.pos="H", message=FALSE, warning=FALSE, paged.print=FALSE}
plot(cars$speed, cars$dist)
```

@@

```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)

mtcars$cyl <- as.factor(mtcars$cyl) # Convert cyl to factor

ggplot(mtcars, aes(x=wt, y=mpg, shape=cyl)) + geom_point() +
  labs(x="Weight (lb/1000)", y = "Miles/(US) gallon", 
       shape="Number of \n Cylinders") + theme_classic()
```

>>>>

---
title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]"
shorttitle: "Writing a reproducible paper with R Markdown and Pagedown"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <camille.landesvatter@mzes.uni-mannheim.de>
date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
download-link: 'Download: https://osf.io/k8jhx'
abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing). Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper". We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and Pagedown. We intend to update this file when we discover more convenient code. You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/.'
# at least one keyword must be supplied
keywords: [open science, transparency, replication, R, markdown, pagedown]
links-to-footnotes: false
paged-footnotes: true
bibliography: references.bib
output: 
  pagedown::jss_paged:
    template: wp_paged.html
    self_contained: true
    css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    csl: american-sociological-association.csl
# uncomment the next line if you only want to knit to HTML and not PDF
knit: pagedown::chrome_print
---

@@

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown. In our view, Pagedown and the underlying technology may replace Latex in the long run. In the near future it should also be possible to produce a PDF with static graph and an equivalent html with interactive graps (see dicsussion [here](https://github.com/rstudio/pagedown/pull/87)).

@@

Below we import an exemplary dataset (you can find `data.csv` in the folder with the other files).

@@

```{r}
dput(data[1:5,]) # here we only take a subset
```

@@

```{r eval=FALSE, include=TRUE}
data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
20.6)), class = "data.frame", row.names = c(NA, -5L))
```

@@

Producing good tables and referencing these tables within a R Markdown document has been a hassle but got much better. Examples that you may use are shown below.

@@

kable(swiss[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)

@@

```{r table-3, echo=TRUE, message=FALSE, warning=FALSE}
library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 title = "Summary: Numeric variables")
```

@@

```{r table-4, echo=TRUE, message=FALSE, warning=FALSE}
# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

library(flextable)
tab_cat <- datasummary_skim(swiss, 
                            type="categorical", 
                            title = "Summary: Categorical variables",
                            output = 'flextable')

@@

```{r fig-1, fig.align="center", fig.cap="Scatterplot of Speed and Distance", fig.pos="H", message=FALSE, warning=FALSE, paged.print=FALSE}
plot(swiss$Catholic, swiss$Fertility)
```

@@

```{r fig-2, fig.align="center", fig.cap="Miles per gallon according to the weight", fig.pos="H", fig.width=6, fig.height=3, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) + geom_point() +
  labs(x="Agriculture", y = "Fertility", 
       shape="Education") + theme_classic()
```
